Bayesian inference of relative fitness on high-throughput pooled competition assays

### 2024-02-26 doc/paper.qmd
working template. Everything looks okay.
prepare paper for submission
add lua filters to split main text and SI text references in the PDF automatically
add new SI section
add kinsler section to paper
Add time scaling section to paper
upload missing document modifications

# Abstract

{{< include _abstract.qmd >}}

@@

# Acknowledgements

{{< include _acknowledgements.qmd >}}

# Supplementary Materials

@@

>>>>

::: {.maintext}

# Abstract {.unnumbered}

The tracking of lineage frequencies via DNA barcode sequencing enables the
quantification of microbial fitness. However, experimental noise coming from
biotic and abiotic sources complicates the computation of a reliable inference.
We present a Bayesian pipeline to infer relative microbial fitness from
high-throughput lineage tracking assays. Our model accounts for multiple sources
of noise and propagates uncertainties throughout all parameters in a systematic
way. Furthermore, using modern variational inference methods based on automatic
differentiation, we are able to scale the inference to a large number of unique
barcodes. We extend this core model to analyze multi-environment assays,
replicate experiments, and barcodes linked to genotypes. On simulations, our
method recovers known parameters within posterior credible intervals. This work
provides a generalizable Bayesian framework to analyze lineage tracking
experiments. The accompanying open-source software library enables the adoption
of principled statistical methods in experimental evolution.

# Author Summary {.unnumbered}

{{< include _author_summary.qmd >}}

@@

# Materials and Methods {.unnumbered}

All synthetic data generation and custom scripts used in this work were stored
using Git version control. Code for analysis and figure generation can be found
on the GitHub repository (https://github.com/mrazomej/bayesian_fitness). The
accompanying software package `BarBay.jl` can be directly installed from the
`Julia` package repository, or via cloning the corresponding GitHub repository
(https://github.com/mrazomej/BarBay.jl)

# Acknowledgements {.unnumbered}

{{< include _acknowledgements.qmd >}}
:::

::: {.sitext}
# Supplementary Materials

@@

<!-- Stochastic growth simulations -->
{{< include _fitseq2_sim.qmd >}}

<!-- Kinsler et al. 2020 -->
{{< include _kinsler.qmd >}}

<!-- Scaling as a function of number of barcodes -->
{{< include _time_scaling.qmd >}}
:::
